Asymptotic Statistics

### 2019-03-06 texts/math/AsymptoticStatistics.tex
.

\begin{align}
	\int_{-\infty}^{\infty}{f(x)dx} = 1 && f(x) >= 0
\end{align}

@@

\begin{align}
	\text{VAR}[X] = E[X^2] - E[X]^2
\end{align}

@@

\begin{align*}
	E[X] &= \sum_{i=1}^{6}{i*P(X=i)}\\
	E[X] &= \sum_{i=1}^{5}{i*P(X=i)} + 6*P(X=6)\\
	E[X] &>= 6*P(X=6)\\
	\frac{E[X]}{6} &>= P(X=6)\\
	P(X=6) &<= \frac{E[X]}{6}
\end{align*}

@@

\begin{align*}
	E[X] &= \int_{0}^{a}{x*f(x)dx} + \int_{a}^{\infty}{x*f(x)dx}\\
	E[X] &= b + \int_{a}^{\infty}{x*f(x)dx} && b > 0\\
	E[X] - b &= \int_{a}^{\infty}{x*f(x)dx}\\
	E[X] &>= \int_{a}^{\infty}{x*f(x)dx} && b > 0\\
\end{align*}

@@

We must realize the in the second integration, $x>=a$ and $f(x) > 0$ which makes the second integration greater than or equal to zero. This allow us to:

\begin{align*}
	\int_{a}^{\infty}{x*f(x)dx} &=\int_{a}^{\infty}{a*f(x)dx}+\int_{a}^{\infty}{(x-a)*f(x)dx}\\
	\int_{a}^{\infty}{x*f(x)dx}	&=\int_{a}^{\infty}{a*f(x)dx}+c\\
	\int_{a}^{\infty}{x*f(x)dx}	- c &=\int_{a}^{\infty}{a*f(x)dx}\\
	\int_{a}^{\infty}{x*f(x)dx}	&>= \int_{a}^{\infty}{a*f(x)dx}\\	
\end{align*}

@@

\begin{align*}
	E[X] &>= \int_{a}^{\infty}{x*f(x)dx} && b > 0\\
	E[X] &>= \int_{a}^{\infty}{a*f(x)dx} && \text{see above}\\
	E[X] &>= a*\int_{a}^{\infty}{f(x)dx}\\
	E[X] &>= a*P(X>a)\\
	\frac{E[X]}{a} &>= P(X>a)\\
\end{align*}
\begin{align}
	P(X>a) &<= \frac{E[X]}{a}	
\end{align}

@@

>>>>

\begin{align}
	\int_{-\infty}^{\infty}{f(x)dx} = 1 && f(x) \ge 0
\end{align}

@@

\begin{align}
	\text{VAR}[X] = E[X^2] - E[X]^2\\
\end{align}

\begin{align}
\text{VAR}[X] = E[(X-E[X])^2]
\end{align}

And we also have the "standard deviation".

\begin{align}
	\text{SD}[X] = \sqrt{\text{VAR}[X]}
\end{align}

@@

\begin{align*}
	E[X] &= \sum_{i=1}^{6}{i*P(X=i)}\\
	E[X] &= \sum_{i=1}^{5}{i*P(X=i)} + 6*P(X=6)\\
	E[X] &\ge 6*P(X=6)\\
	\frac{E[X]}{6} &\ge P(X=6)\\
	P(X=6) &\le \frac{E[X]}{6}
\end{align*}

@@

\begin{align*}
	E[X] &= \int_{0}^{a}{x*f(x)dx} + \int_{a}^{\infty}{x*f(x)dx}\\
	E[X] &= b + \int_{a}^{\infty}{x*f(x)dx} && b > 0\\
	E[X] - b &= \int_{a}^{\infty}{x*f(x)dx}\\
	E[X] &\ge \int_{a}^{\infty}{x*f(x)dx} && b > 0\\
\end{align*}

@@

We must realize the in the second integration, $x \ge a$ and $f(x) > 0$ which makes the second integration greater than or equal to zero. This allow us to:

\begin{align*}
	\int_{a}^{\infty}{x*f(x)dx} &=\int_{a}^{\infty}{a*f(x)dx}+\int_{a}^{\infty}{(x-a)*f(x)dx}\\
	\int_{a}^{\infty}{x*f(x)dx}	&=\int_{a}^{\infty}{a*f(x)dx}+c\\
	\int_{a}^{\infty}{x*f(x)dx}	- c &=\int_{a}^{\infty}{a*f(x)dx}\\
	\int_{a}^{\infty}{x*f(x)dx}	&\ge \int_{a}^{\infty}{a*f(x)dx}\\	
\end{align*}

@@

\begin{align*}
	E[X] &\ge \int_{a}^{\infty}{x*f(x)dx} && b > 0\\
	E[X] &\ge \int_{a}^{\infty}{a*f(x)dx} && \text{see above}\\
	E[X] &\ge a*\int_{a}^{\infty}{f(x)dx}\\
	E[X] &\ge a*P(X \ge a)\\
	\frac{E[X]}{a} &\ge P(X \ge a)\\
\end{align*}
\begin{align}
	P(X \ge a) &\le \frac{E[X]}{a}	
\end{align}

@@

One of the interesting aspects of the "Markov Inequality" is that it only contains two assumptions and both are very generic. It only depends on X being non-negative and a being nonnegative. We know that a nonnegative function, when sifted on the X-Axis is still nonnegative. So $X-a$ is still non-negative and the "Markov Inequality" still apllies.

\begin{align*}
P((X-b) \ge a) &\le \frac{E[X-b]}{a}	
\end{align*}

Another "transformation" that keeps the non-negativiness of the function is squaring its value.

\begin{align*}
P((X-b)^2 \ge a^2) &\le \frac{E[(X-b)^2]}{a^2}	
\end{align*}

But we do not have any restriction on $b$ here, and we can easily choose the "expected value" of $X$.

\begin{align*}
P((X-E[X])^2 \ge a^2) &\le \frac{E[(X-E[X])^2]}{a^2}
\end{align*}

Which gives us the $\text{VAR}[X]$. Although this inequality is often writen in a different form.

\begin{align*}
P((X-E[X])^2 \ge a^2) &\le \frac{\text{VAR}[X]}{a^2}\\	
\end{align*}

But this is still not the most "famous" way this formula is written, and the reason is going to be appear in future chapters. To derive the most famous form we are going to apply some transformations only to the left side, because they will not change the boundness of the inequality.

\begin{align*}
P(\sqrt{(X-E[X])^2}\ge\sqrt{a}) &\le \frac{\text{VAR}[X]}{a^2}\\
P(|X-E[X]| \ge a) &\le \frac{\text{VAR}[X]}{a^2} && \text{ no modulus on a because } a>0
\end{align*}

\begin{align}
	P(|X-E[X]| \ge a) &\le \frac{\text{VAR}[X]}{a^2}
\end{align}

Which is quite fascinating, because it gives you a know bound when you know the distribution "expected value" and "variance". This formula will demonstrate itself very powerfull and useful in the study of sampling and convergence.

We can also interpret that it is "harder", less probable, of being $n$ "standard deviations" from the "expected value". What is very intuitive, if you stop to think about it. What you can not do, is forget the "standard deviation" and say that it is "harder" to be far from the "expected value". This is not true. Without the "standard deviation", or "variance" this phrase, does not make any send. Stop to think of a distribution with "infinite" "variance". It is very "easy" to be "infinitilly" far from the "expected value".

\section{Markov's and Chebyshev's Inequalities}

If you search you will probably find another formula for both inequalities. They are all equivalent, off course.

\begin{align*}
	P(X \ge a) \le \frac{E[X]}{a}\\	
	P(X \ge a*E[X]) \le \frac{E[X]}{a*E[X]} && E[X] > 0
\end{align*}
\begin{align}
	P(X \ge a*E[X]) \le \frac{1}{a}
\end{align}

\begin{align*}
P((X-E[X])^2 \ge a^2) &\le \frac{E[(X-E[X])^2]}{a^2}\\
P((X-E[X])^2 \ge a^2*\text{VAR}[X]) &\le \frac{E[(X-E[X])^2]}{a^2*\text{VAR}[X]}\\
P((X-E[X])^2 \ge a^2*\text{VAR}[X]) &\le \frac{\text{VAR}[X]}{a^2*\text{VAR}[X]}\\
P((X-E[X])^2 \ge a^2*\text{VAR}[X]) &\le \frac{1}{a^2}\\
P(\sqrt{(X-E[X])^2}\ge\sqrt{a^2*\text{VAR}[X]}) &\le \frac{1}{a^2}\\
\end{align*}
\begin{align}
P(|(X-E[X])| \ge a*\text{SD}[X]) &\le \frac{1}{a^2}
\end{align}

\section{68-95-99.7 rule}

\section{Summary}

\subsection{Markov's Inequality}

\begin{align}
P(X \ge a) &\le \frac{E[X]}{a}
\end{align}

\begin{align}
P(X \ge aE[X]) &\le \frac{1}{a}
\end{align}

\subsection{Chebyshev's Inequality}

\begin{align}
P(|X - E[X] \ge a) &\le \frac{\text{VAR}[X]}{a^2}
\end{align}

\begin{align}
P(|X - E[X] \ge a*\text{SD}[X]) &\le \frac{1}{a^2}
\end{align}
